---
title: "MS Fees 2019"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Adrian D. Garcia"
date: "June 2020"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse) #For Data Analysis
library(lubridate) #For working with dates
library(DT) #For visualizing tables
```

First, I pull in the data from Morningstar Direct (combined using "Combine_Files.r"). There are 25,654 rows and 96 columns.
Data was pulled from Morningstar Direct on June 12, 2020.

```{r load, message= FALSE, warning= FALSE}


Full <- read_csv("DIRECT/MS_Fee_Report_061220.csv",
                 guess_max = 5000)

```

I'm making the following adjustment to the data:

-Calculate the average assets for 2017, 2018 and 2019.
-Calculate the percent change in net expense ratios from 2018 to 2019
-Multipy the annual average asset figures by the expense ratio (to use later for asset-weighted calculations)

```{r adjustment, echo=FALSE}
Adjusted_Data <- Full %>%
  mutate(
    Net_Asst_2017_AVG = rowMeans(Full[,53:64], na.rm=TRUE),
    Net_Asst_2018_AVG = rowMeans(Full[,65:76], na.rm=TRUE),
    Net_Asst_2019_AVG = rowMeans(Full[,77:88], na.rm=TRUE),
    Pts_2018_2019 = `Annual Report \nNet Expense \nRatio \nYear2019` - `Annual Report \nNet Expense \nRatio \nYear2018`,
    Change_2018_2019 = ((`Annual Report \nNet Expense \nRatio \nYear2019`/100)-(`Annual Report \nNet Expense \nRatio \nYear2018`/100))
    /(`Annual Report \nNet Expense \nRatio \nYear2018`/100)*100,
    Asst_By_Fee_2017 = `Annual Report \nNet Expense \nRatio \nYear2017` * Net_Asst_2017_AVG,
    Asst_By_Fee_2018 = `Annual Report \nNet Expense \nRatio \nYear2018` * Net_Asst_2018_AVG,
    Asst_By_Fee_2019 = `Annual Report \nNet Expense \nRatio \nYear2019` * Net_Asst_2019_AVG,
  )
```

Here I want to look how deep firms slashed pay ratios from 2018 to 2019.
First, remove rows with an expense ratio of zero in 2018 because I can't calculate changes with that base. I also filter out expense ratios listed as NA (blank) for 2018 and 2019.

Now I need to put the percent change in fees from 2018 to 2019 into buckets. I did that with the cut() function. Here's a for personal reference:
https://rpubs.com/pierrelafortune/cutdocumentation

```{r cuts_by_shareclass, message = FALSE}

Category <- Adjusted_Data %>% 
  select(
    1:20,
    86:96
  ) %>%
  filter(
    `Annual Report \nNet Expense \nRatio \nYear2018` != 0 &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2018`) &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`)
  ) %>% 
  mutate(
    Pct_FeeChange = 
      cut(Change_2018_2019,
          breaks = c(-Inf,-50,-25,-10,-1,-0.01,0,Inf),
          labels = c("< -50", "-25 - -50", "-10 - -25", "-1 - 10","0 - -1", "0", "> 0"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Summary_Category <- Category %>% 
  group_by(Pct_FeeChange) %>% 
  summarise(count = n())

ggplot(data = Summary_Category, mapping = aes(x = Pct_FeeChange, y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = 0)

```

Here I am looking at the fund families that with the lowest and highest overall fees across active products. I start by filtering out index funds and removing fees that are NA (blank) for 2019. Next I group the data by firm, sum their adjusted assets, sum their net assets and divide the total adjusted data by the total net asssets to calculate asset-weighted fees for each fund shop. I select the top and bottom 15 fund families based on the results.

```{r active_fee, message=FALSE}

ActiveFunds <- Adjusted_Data %>% 
  filter(
    `Index \nFund` != "Yes" &
            !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`)

  )

Active_Fees <- ActiveFunds %>% 
  group_by(`Firm Name`) %>% 
  summarise(
    Sum_Assts_Wted = sum(Asst_By_Fee_2019, na.rm = TRUE),
    Sum_Assets = sum(Net_Asst_2019_AVG, na.rm = TRUE),
    Fee_Wted = Sum_Assts_Wted/Sum_Assets
  )

Active_HighFees <- Active_Fees %>% 
  arrange(desc(Fee_Wted)) %>% 
  top_n(15, Fee_Wted)

Active_LowFees <- Active_Fees %>% 
  arrange(Fee_Wted) %>% 
  top_n(-15, Fee_Wted)

ggplot(data = Active_HighFees, mapping = aes(x = reorder(`Firm Name`, Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

ggplot(data = Active_LowFees, mapping = aes(x = reorder(`Firm Name`, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

```

Here I am looking at the fund families that with the lowest and highest overall fees across passive products. I start by filtering out non-index funds and removing fees that are NA (blank) for 2019. Next I group the data by firm, sum their adjusted assets, sum their net assets and divide the total adjusted data by the total net asssets to calculate asset-weighted fees for each fund shop. I select the top and bottom 15 fund families based on the results.

```{r passive_fee, message=FALSE}

PassiveFunds <- Adjusted_Data %>% 
  filter(
    `Index \nFund` != "No" &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`)
  )

Passive_Fees <- PassiveFunds %>% 
  group_by(`Firm Name`) %>% 
  summarise(
    Sum_Assts_Wted = sum(Asst_By_Fee_2019, na.rm = TRUE),
    Sum_Assets = sum(Net_Asst_2019_AVG, na.rm = TRUE),
    Fee_Wted = Sum_Assts_Wted/Sum_Assets
  )

Passive_HighFees <- Passive_Fees %>% 
  arrange(desc(Fee_Wted)) %>% 
  top_n(15)

Passive_LowFees <- Passive_Fees %>% 
  arrange(Fee_Wted) %>% 
  top_n(-15)

ggplot(data = Passive_HighFees, mapping = aes(x = reorder(`Firm Name`, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

ggplot(data = Passive_LowFees, mapping = aes(x = reorder(`Firm Name`, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

```

Here I am looking at how many share classes had fees that were cut and hiked as well as saw no change. I start by filtering out fees that are NA (blank) for 2019 and 2018. I also remove fees with a zero expense ratio for 2018 because we can't calculate change with that base. Next I create a category (Pct_FeeChange) that puts the percent changes (continous) in buckets (categorical). After that I just count the buckets.

```{r cuts_by_funds, message = FALSE}

Shareclass <- Adjusted_Data %>% 
  filter(
    `Annual Report \nNet Expense \nRatio \nYear2018` != 0 &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2018`) &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`)
  ) %>% 
  mutate(
    Pct_FeeChange = 
      cut(Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Summary_Shareclass <- Shareclass %>% 
  group_by(`Firm Name`,
           Pct_FeeChange) %>% 
  summarise(count = n())

ggplot(data = Summary_Shareclass, mapping = aes(x = Pct_FeeChange, y = count)) +
  geom_bar(stat = "identity")

```

 Here I am looking at how many Passive share classes had fees that were cut and hiked as well as saw no change. I start by filtering out non-index funds and fees that are NA (blank) for 2019 and 2018. I also remove fees with a zero expense ratio for 2018 because we can't calculate change with that base. Next I create a category (Pct_FeeChange) that puts the percent changes (continous) in buckets (categorical). After that I just count the buckets.

```{r cuts_by_passive_funds, message = FALSE}

Passive_Shareclass <- Adjusted_Data %>% 
  filter(
    `Annual Report \nNet Expense \nRatio \nYear2018` != 0 &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2018`) &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`) &
      `Index \nFund` == "Yes"
  ) %>% 
  mutate(
    Pct_FeeChange = 
      cut(Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Passive_Summary_Shareclass <- Passive_Shareclass %>% 
  group_by(`Firm Name`,
           Pct_FeeChange) %>% 
  summarise(count = n()) %>% 
  group_by(`Firm Name`) %>% 
  mutate(
    pct = count/sum(count)*100)

Passive_Summary_Shareclass %>% 
  filter(Pct_FeeChange == "Cut") %>% 
  arrange(desc(count)) %>% 
  ungroup() %>% 
  top_n(15, count) %>% 
ggplot(mapping = aes(x = reorder(`Firm Name`, count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = count), vjust = 0)

```

tk

```{r cuts_by_active_funds, message = FALSE}

Active_Shareclass <- Adjusted_Data %>% 
  filter(
    `Annual Report \nNet Expense \nRatio \nYear2018` != 0 &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2018`) &
      !is.na(`Annual Report \nNet Expense \nRatio \nYear2019`) &
      `Index \nFund` == "Yes"
  ) %>% 
  mutate(
    Pct_FeeChange = 
      cut(Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Active_Summary_Shareclass <- Active_Shareclass %>% 
  group_by(`Firm Name`,
           Pct_FeeChange) %>% 
  summarise(count = n()) %>% 
  group_by(`Firm Name`) %>% 
  mutate(
    pct = count/sum(count)*100)

Active_Summary_Shareclass %>% 
  filter(Pct_FeeChange == "Cut") %>% 
  arrange(desc(count)) %>% 
  ungroup() %>% 
  top_n(15, count) %>% 
ggplot(mapping = aes(x = reorder(`Firm Name`, count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = count), vjust = 0)

```